---
title: "Extracting Information about Treatments"
author: "TANG"
date: "2024-02-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Extracting Information about Treatments

# Control / Nesting (Exercise 8.5)
Re-analyse the data of Exercise 6.3 using new factors (and information supplied
in file scab.dat) to answer the following questions: Does the addition of sulphur
affect the level of scab? Does either of the rate or timing of application affect the
level of scab? Do these two factors act independently? (We re-visit these data in
Exercise 11.4.)

## Set up
```{r}
# Read data & assign factors
scab <- read.table('scab.dat', sep="", header=TRUE)
head(scab)
summary(scab)
```
Calculation of Sulphur Parental Effects, and Nested Sulphur.Time Effects
```{r}
# Convert Block and Plot to factors
scab$Plot <- as.factor(scab$Plot)
scab$Row <- as.factor(scab$Row)
scab$Col <- as.factor(scab$Col)

# One-way ANOVA with blocking and nested treatment structure (multi-stratum)
scab.msaov <- aov(Scab ~ Sulphur/Timing + Error(Plot/Row/Col), data=scab)
summary(scab.msaov)

# Print effects for Type and Fungicide nested witin Type
model.tables(scab.msaov, type="means")

# Print effects for single treatment factor
scab.msaov2 <- aov(Scab ~ Sulphur + Error(Row/Col), data=scab)
model.tables(scab.msaov2, type="effects")
```


# Polynomial contrasts (Exercise 8.8)
Consider the data from the calcium pot trial of Example 4.1 (Table 4.1 and file
calcium.dat). In this trial, the treatments A, B, C and D were concentrations of
calcium in the soil, measured as relative concentrations of 1, 5, 10 and 20, respectively.

Re-analyse these data using polynomial contrasts. Which low-order polynomial
provides the best fit to these data?

## Set up
```{r}
# load external packages - available from CRAN
library(ggplot2)

# Read data & assign factors
calcium <- read.table('calcium.dat', sep="", header=TRUE)
head(calcium)
summary(calcium)

# Plot data
qplot(data=calcium, y=Length, x=Calcium)

# define vCalcium as a variate of concentrations
calcium$vCalcium <- rep(0,length(calcium$Calcium))
calcium$vCalcium[calcium$Calcium=="A"] <- 1
calcium$vCalcium[calcium$Calcium=="B"] <- 5
calcium$vCalcium[calcium$Calcium=="C"] <- 10
calcium$vCalcium[calcium$Calcium=="D"] <- 20
```
## linear regression
```{r}
# Baseline ANOVA
calcium.aov <- lm(Length ~ vCalcium, data=calcium)
anova(calcium.aov)
summary(calcium.aov)

# Get coefficients
coef <- coef(calcium.aov)
coef

# Calculate predictions from quadratic polynomial
c <- c(1:20)
fit <- coef[1] + coef[2]*c 

# Plot predictions with data and fitted means
# Set up basic plot using model object
plot.obs <- ggplot(data=calcium, aes(y=Length, x=vCalcium) ) 
# Plot observations and fitted line
plot.obs + geom_point() + 
  geom_point(aes(y=fitted(calcium.aov)),colour="blue",size=2,shape=2) +
  geom_line(aes(y=fit,x=c),colour = "red") +
  ggtitle("Observed data with predicted means and linear")
```

## Polynomial contrasts
```{r}
# To get coefficients with respect to un-transformed variable, refit quadratic 
calcium2.aov <- lm(Length ~ 1 + vCalcium + I(vCalcium^2), data=calcium)
anova(calcium2.aov)
summary(calcium2.aov)

# Get coefficients
coef2 <- coef(calcium2.aov)
coef2

# Calculate predictions from quadratic polynomial
c <- c(1:20)
fit2 <- coef2[1] + coef2[2]*c + coef2[3]*c*c

# Plot predictions with data and fitted means
# Set up basic plot using model object
plot.obs <- ggplot(data=calcium, aes(y=Length, x=vCalcium) ) 
# Plot observations and fitted line
plot.obs + geom_point() + 
  geom_point(aes(y=fitted(calcium.aov)),colour="blue",size=2,shape=2) +
  geom_line(aes(y=fit2,x=c),colour ="darkgreen") +
  ggtitle("Observed data with predicted means and quadratic polynomial")
```

# End of file


